---
title: "Introduction to CoVVSURF"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Put the title of your vignette here}
  %\VignetteEngine{knitr::rmarkdown}
  \usepackage[utf8]{inputenc}
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,eval=FALSE)
```

This document is a short introduction to the R package CoVVSURF which combines clustering of variables and feature selection using random forest. The procedure CoV/VSURF is a two steps methodology for dimension reduction and variable selection in the context of supervised classifcation (which can also be applied for regression problems). First, one eliminates
redundancy using clustering of variables, based on the R package __ClustOfVar__. This clustering approach, denoted by __CoV__ hereafter, allows one to deal with both numerical and categoricalcvariables. The clustering of variables groups together highly correlated variables and provides for each group (cluster)ca synthetic variable which is a numerical variable summarizing the variables within a cluster. The main advantage of
this step is to eliminate redundancy and to keep all the variables together in a cluster during the rest of the analysis. Moreover it reduces the dimension of the data by replacing the p original variables by $K$ synthetic variables (where $K$ denotes the selected number of clusters). Note that this clustering of variables approach does not require définition
of a priori groups of variables as in group lasso or sparse group lasso approaches. In addition, the reduction of dimension provides $K$ synthetic variables which only use the variables within a cluster, unlike
the principal components in principal component analysis (PCA). Hence, in __CoV__, an original variable takes action in the construction of a unique synthetic variable, which make the interpretation easier.
Secondly, the most important synthetic variables obtained in the first step are selected using a procedure based on random forests (RF) implemented in the R package __VSURF__. This variable selection procedure, denoted __VSURF__ hereafter, is applied to the reduced dataset consisting of
the $n$ observations described with the $K$ synthetic variables. Thus a list of selected synthetic variables (i.e. a list of clusters of variables) is obtained and the prediction for new observations can be done with a predictor built on these selected synthetic variables

Details are given in [https://arxiv.org/abs/1608.06740](https://arxiv.org/abs/1608.06740).

# Installation

```{r}
devtools::install_github("chavent/PCAmixdata")
devtools::install_github("robingenuer/CoVVSURF")
```


# The classification context

```{r}
library(CoVVSURF)
help(package=CoVVSURF)
```


## Simulated dataset
The data are simulated from an underlying classification model where 120 explonatory variables (80 quantitative and 40 qualitative with underlying informative and non informative groups) explain a binary response variable. The function __simu_classif__ generates $n$ observations of the $p=120$ explonatory variables in a matrix $X$ and of the binary reponse in a vector $y$.


```{r}
train <- simu_classif(n=60,seed=10) # simulated training data set
train$X #explonatory data matrix
train$y #binary variable to predict
```

##Clustering of variables
The function __CoV__ builds a dendrogram of variables using the R package __ClustOfVar__.

\bigskip

```{r}
treecov <- CoV(train$X) # dendrogram of the 120 variables with ClustOfVar
plot(treecov)
```

##The CoV/VSURF procedure for classification
The functions __covsurf__ and __predict.covsurf__ implement the CoV/VSURF procedure. The input of the procedure is a dataset $(X,y)$ and the goal is to select groups of informative variables in $X$ to predict $y$.  

In the first step the function __covsurf__ selects groups of informative variables in the following way: 
  
  1. Build the dendrogram of variables  with __CoV__ 
  2. Let $kval$ be the set of number of groups to try. For each $k \in kval$, cut __CoV__ tree (dendrogram) in $k$ clusters, train a Random Forest (RF) with the $k$ synthetic variables $f^1,...,f^k$ as predictors and $y$ as output variable and compute the mean OOB error rates.
  3. Choose the optimal number $kopt$ of clusters, which leads to the minimum mean OOB error rate. Cut __CoV__ tree in $kopt$ clusters.
  4. Perform VSURF with the $kopt$ synthetic variables $f^1,...,f^{kopt}$ 
as predictors and $y$ as output variable. Denote by $m \leq kopt$ the number of selected informative synthetic variables (corresponding to the interpretation set of VSURF).

\bigskip
Parallel computing with several cores can be used to speed the execution.
\bigskip
```{r}
kval <- c(2:15, seq(from = 20, to = ncol(train$X), by = 10))
res <- covsurf(train$X, train$y, kval) 
#or to seed the execution :
# if the tree has already been built
res <- covsurf(train$X, train$y, tree=treecov,kval) 
# and if you  can use several cores
res <- covsurf(train$X, train$y, tree=treecov,kval,ncores=3) 
names(res)
plot(res)
res$kopt # the partition in 10 cluster of the dendrogram is selected
res$vsel # 5 synthetic variables are selected by VSURF
res$csel # the corresponding 5 selected groups of variables 
```

In the seconde step the function  __predict.covsurf__ gives the prediction of a set of new observations $x$ in the following way:

  1. Train a random forest, $\hat{f}$, on the dataset consisting of the $m$ selected synthetic variables and $y$.
  2. Compute the scores of $x$ on the $m$ selected synthetic variables and predict its class label using $\hat{f}$.

\bigskip
```{r}
test <- simu_classif(n=100,seed=20) # simulated test data set
```

```{r}
pred <- predict(res,test$X)
sum(pred==test$y)/length(test$y) # True classification rate
```

# The regression context


```{r}
library(CoVVSURF)
help(package=CoVVSURF)
```

## Simulated dataset
The data are simulated from an underlying classification model where 120 explonatory variables (80 quantitative and 40 qualitative with underlying informative and non informative groups) explain a numeric response variable. The function __simu_reg__ generates $n$ observations of the $p=120$ explonatory variables in a matrix $X$ and of the numeric reponse in a vector $y$.

```{r}
train <- simu_reg(n=60,seed=10) # simulated training data set
train$X #explonatory data matrix
train$y #numeric variable to predict
```

##Clustering of variables
The function __CoV__ builds a dendrogram of variables using the R package __ClustOfVar__.

\bigskip

```{r}
treecov <- CoV(train$X) # dendrogram of the 120 variables with ClustOfVar
plot(tree)
```

##The CoV/VSURF procedure for regression
The functions __covsurf__ and __predict.covsurf__ implement the CoV/VSURF procedure. The procedure is similar to that presented previously in the context of the classification with the OOB mean square errors (sum of squared residuals divided by $n$) replacing the OOB error rate.

\bigskip
In the first step the function __covsurf__ selects groups of informative variables. Parallel computing with several cores can be used to speed the execution.

```{r}
kval <- c(2:15, seq(from = 20, to = ncol(train$X), by = 10))
res <- covsurf(train$X, train$y, kval)
#or if the tree has already been built
res <- covsurf(train$X, train$y, tree=treecov,kval) #be patient...
#or if several cores
res <- covsurf(train$X, train$y, tree=treecov,kval,ncores=3) 
names(res)
plot(res)
res$kopt # the partition in 9 cluster of the dendrogram is selected
res$vsel # 6 synthetic variables are selected by VSURF
res$csel # the corresponding 6 selected groups of variables 
```

In the seconde step the function  __predict.covsurf__ gives the prediction of a set of new observations $x$.

\bigskip
```{r}
test <- simu_reg(n=100,seed=20) # simulated test data set
```

```{r}
pred <- predict(res,test$X)
sum(pred-test$y)^2/length(test$y) # mean square error
```
